---
title: "Oncology Clinical Trial Analysis"
author: "Najiya K"
output: html_document
---

# 🧬 Project Overview

This project evaluates the effectiveness and safety of a treatment compared to placebo using survival analysis and adverse event data.

---

# 📦 Load Packages and Data

```{r}
library(haven)
library(dplyr)
library(survival)

adsl <- read_sas("adsl.sas7bdat")
adtte <- read_sas("adtte.sas7bdat")
adae <- read_sas("adae.sas7bdat")
```

---

# 🧪 Overall Survival (OS)

## Prepare Dataset

```{r}
adtte_os <- adtte %>%
  filter(PARAMCD == "OS") %>%
  left_join(adsl, by = "USUBJID")
```

---

## Kaplan-Meier Curve

```{r fig.width=6, fig.height=5}
km_fit <- survfit(Surv(AVAL, CNSR == 0) ~ TRT01P, data = adtte_os)

plot(km_fit, col = c("blue", "red"), lwd = 2,
     xlab = "Time", ylab = "Survival Probability")

legend("topright",
       legend = c("Placebo", "Treatment"),
       col = c("blue", "red"),
       lwd = 2)
```

### 🧠 Insight
The treatment group shows improved survival compared to placebo.

---

## Log-Rank Test

```{r}
logrank_test <- survdiff(Surv(AVAL, CNSR == 0) ~ TRT01P, data = adtte_os)
logrank_test
```

### 🧠 Insight
The difference between groups is statistically significant.

---

## Cox Model (Hazard Ratio)

```{r}
cox_os <- coxph(Surv(AVAL, CNSR == 0) ~ TRT01P, data = adtte_os)
summary(cox_os)
```

### 🧠 Insight
The treatment reduces the risk of death compared to placebo.

---

# 🔬 Progression-Free Survival (PFS)

## Prepare Dataset

```{r}
adtte_pfs <- adtte %>%
  filter(PARAMCD == "PF") %>%
  left_join(adsl, by = "USUBJID")
```

---

## Kaplan-Meier Curve

```{r fig.width=6, fig.height=5}
km_fit_pfs <- survfit(Surv(AVAL, CNSR == 0) ~ TRT01P, data = adtte_pfs)

plot(km_fit_pfs,
     col = c("blue", "red"),
     lwd = 2,
     xlab = "Time",
     ylab = "Progression-Free Survival Probability")

legend("topright",
       legend = c("Placebo", "Treatment"),
       col = c("blue", "red"),
       lwd = 2)
```

---

## Cox Model (Adjusted)

```{r}
cox_pfs <- coxph(Surv(AVAL, CNSR == 0) ~ TRT01P + AGE + SEX, data = adtte_pfs)
summary(cox_pfs)
```

### 🧠 Insight
No statistically significant improvement in progression-free survival was observed.

---

# 💊 Safety Analysis (ADAE)

## Prepare Dataset

```{r}
adae2 <- adae %>%
  left_join(adsl, by = "USUBJID")
```

---

## Total Adverse Events

```{r}
adae_summary <- adae2 %>%
  group_by(TRT01A) %>%
  summarise(
    total_AE = n(),
    patients = n_distinct(USUBJID)
  )

adae_summary
```

### 🧠 Insight
Adverse events are similar between treatment and placebo groups.

---

## Severity Distribution

```{r}
ae_severity <- adae2 %>%
  group_by(TRT01A, AESEV) %>%
  summarise(count = n())

ae_severity
```

### 🧠 Insight
No increase in severe adverse events in the treatment group.

---

## Common Adverse Events

```{r}
common_ae <- adae2 %>%
  group_by(TRT01A, AEDECOD) %>%
  summarise(count = n()) %>%
  arrange(TRT01A, desc(count)) %>%
  group_by(TRT01A) %>%
  slice_head(n = 5)

common_ae
```

### 🧠 Insight
Common adverse events include headache, fatigue, nausea, and neutropenia.

---

# ⚖️ Final Conclusion

The treatment demonstrated a statistically significant improvement in overall survival compared to placebo, while no significant difference was observed in progression-free survival. Safety analysis showed a comparable incidence and severity of adverse events between treatment and placebo groups, with no increase in severe adverse events. Overall, the benefit–risk profile of the treatment appears favorable.